A Global Genome-wide Scan with Optimal Cutoff Mining for Emerging Biomarkers in Head and Neck Squamous Cell Carcinoma

### 2021-02-15 main.tex
Updates from Overleaf
next on \subsection{Research limitation}

\newenvironment{MyColorPar}[1]{%
    \leavevmode\color{#1}\ignorespaces%
}{%
}%
%\usepackage[upplemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
%...
%\usepackage{graphicx}
%\DeclareGraphicsExtensions{.pdf,.png} % for high-resolution PDF image
%\usepackage{hyperref}
%\usepackage{url}

@@

%latex https://www.latex-tutorial.com/symbols/greek-alphabet/
\subsection*{Linear regression in biomarker discovery}
Linear regression is useful as a concept in biomarker discovery.
A linear regression equation that relates a Y-variable to n X-variables is written as%\\[1cm]
\par
%Highlights
% https://quicklatex.com
\begin{flushleft}
% equation
%\begin{math}

@@

\subsection{A introduction of TCGA}
%\begin{MyColorPar}{red}
%%%%%% revision1 added [TCGA] [2021/02/02]
A large-scale cancer database, aggregating many $X_1 ... X_n$ features, is necessary to facilitate the biomarker discovery.
The Cancer Genome Atlas (TCGA) project\cite{Weinstein2013} has been developed since 2005 and supervised by the National Cancer Institute's (NCI) Center for Cancer Genomics and the National Human Genome Research Institute (NHGRI), funded by the US government.
TCGA represents comprehensive genomics and clinic data from 84,392 patients among 33 major cancer types (data release 27.0 - Oct 29, 2020, available at https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga/studied-cancers).
%The Cancer Genome Atlas (TCGA): The TCGA consortium, which is a National Institute of Health (NIH) initiative, makes publicly available molecular and clinical information for more than 33 types of human cancers including exome (variant analysis), single nucleotide polymorphism (SNP), DNA methylation, transcriptome (mRNA), microRNA (miRNA) and proteome.
TCGA and other collaborated \acrfull{gdac} generated and analyzed DNA (mutation, copy number variation, methylation, simple nucleotide polymorphism, SNP), RNA (microarray, RNA-Seq, microRNA), and protein (reverse protein phased array) data derived from biospecimen. Sample types available at TCGA are primary solid tumors, recurrent solid tumors, blood-derived normal and tumor, metastatic, and solid tissue normal.

@@

% to summarize the advantages of its use.
The highlights of the advantages of applying the TCGA data for cancer biomarker identification include:
\begin{outline}
\1  To the best of our knowledge, the TCGA database is the largest collection (in both cancer types and cohort size) of comprehensive genomics with survival data available in the field of cancer research so far.  
There are many physical and social features ($X_1 ... X_n$) of patients available for survival modeling.
The whole-genome sequencing data has been harmonized across all GDACs.  Many databases adopt the essential demographic data from TCGA since it has comprehensive clinical features, such as exposure to alcohol, asbestos, radioactive radon, tobacco smoking, or cigarettes.
\1  TCGA has a remarkable advantage for computational and life scientists who study cancer since useful web-tools and API are ready to analyze and visualize TCGA data. It might be getting help quickly from the research community for trouble-shooting.
\1  Many achievements of diagnoses, treatments, and prevention with the TCGA data have already been published and keep growing.\cite{Tomczak2015}
\end{outline}

%\end{MyColorPar}
%%%%%% end of additional revision1
\par
In summary, identifying predictive biomarkers for selecting standard-of-care or advanced systemic therapy\cite{Cristina2019} in \acrshort{hnscc} is crucial. 
However, there are three challenges of biomarker discovery from survival analysis.
First, although \acrshort{tcga} genomics data was harmonized, there is unclean data, including null expressed genes, over 50\% of the cohort, should be manually investigated and cleaned.
Second, we need to find a way to determine candidates from the expression level of 20,500 human protein-coding genes\cite{Clamp2007}. Usually, the investigators should get the rationale or revelation of the genes of interest on a specific cancer type. They should upload those genes manually onto bioinformatics tools, such as SurvExpress (http://bioinformatica.mty.itesm.mx:8080/Biomatec/ \newline
SurvivaX.jsp\cite{Aguirre-Gamboa2013}, which has been lost since Oct/2019 and currently out of funds), and analyze the cohort (e.x. \acrshort{tcga}). After downloading the survival results, they could curate plots and tables carefully.
It is not possible to scan the whole human protein-coding genome in this way.
Third, we need to find an optimal cutpoint of that \acrshort{rna} expression data to maximize candidate mining coverage. The above mentioned web-tools might set a cutpoint at the median, 1/4 quantile, or 3/4 quantile for subsequent analyses. There are several visualization software or R packages which deal with cutoff determination, such as Prognoscan\cite{Mizuno2009a}, Cutoff Finder\cite{Budczies2012}, Findcut\cite{Chang2017a}, Human protein atlas\cite{Uhlen2017}, OptimalCutpoints\cite{Cristina2019},  cutpointr (available at https://github.com/thie1e/ \newline
cutpointr), and cutoffR (available at https://cran.r-project.org/web/
packages/cutoffR). However, non of them could combine the scanning of the protein-coding genes and cutoff optimization programmatically.

@@

% TCGAbiolinks
%alcohol consumption
Besides ethnicity, age, gender, the \acrshort{tnm} stage, radiation therapy, chemotherapy, and targeted therapy, the comprehensive adverse features of prognosis in \acrshort{hnscc} should include tobacco exposure, \acrshort{egfr} amplification, \acrfull{hpv} status, positive/close surgical margin ($<5 mm$), \acrfull{ene}, \acrfull{lvsi}, \acrfull{pni}, \acrfull{doi} ($>5 mm$), as well as metastatic \acrfull{lnd}\cite{Cheraghlou2018}, and \acrfull{wpoi5}, which was defined as tumor dispersion $>= 1 mm$ between tumor satellites or positive \acrshort{pni}/\acrshort{lvsi}\cite{Amin2017}.
The features of \acrshort{doi}, \acrshort{lnd}, and tumor dispersion were not available on the \acrshort{tcga} dataset. The Brandwein-Gensler's risk model (lymphocytic host response, \acrshort{wpoi5}, and \acrshort{pni})\cite{Brandwein-Gensler2010}\cite{Sinha2018} was suggested to be routinely performed on pathological examination. In previous reports of \acrshort{hnscc}, the loco-regional failure will be high when the initial frozen section has a positive/close surgical margin, and even the final margin revision revealed negative\cite{Bulbul2019b}.
According to Table \ref{table:table2} and Table \ref{table:table4} in our study, the positive surgical margin yielded a hazard ratio greater than 1.6 to influence on patient's \acrshort{os}.
It was suggested by authors \cite{Scholl1986}\cite{Sutton2003}\cite{Shaw2004}\cite{Guillemaud2010a}\cite{Patel2010}\cite{Kuriakose2017}\cite{Shapiro2017}\cite{Saidak2018}\cite{Miguelanez-Medran2019}\cite{Saidak2019} that the reason of positive/close surgical margin is possibly due to tumor aggressiveness or dispersion (WPOI-5) instead of iatrogenic reason of surgery. The surgical margin status is suggested as an independent surrogate for tumor dispersion, and important in the \acrshort{hnscc} study. Thus, in the current study, we selected the common clinicopathological features, including gender, age, clinical T, clinical N, clinical M, surgical margin status, and tobacco exposure in the biomarker discovery for adjustment of confounders (details description at Materials and Methods section).

In our previous work, altered glucose metabolism (e.k.a. the Warburg effect\cite{Warburg1956}) promotes the progression of \acrshort{hnscc}, which is partially attributed to the \acrfull{slc2a4} (or \acrfull{glut4}) and \acrfull{trim24} pathway\cite{Chang2017b}. In this study, LOC148709 (a \acrfull{lncrna}) was suggested as a biomarker of \acrshort{hnscc} (see Table \ref{table:table3}).
%on page \pageref{table:table3}). I
It was also found to has a contribution to the Warburg effect on esophagus cancer\cite{Liu2019}.

Although we combined the power of genome-wide scanning and an optimal cutoff finder for Kaplan-Meier survival analysis, the study has some limitations.
We are aware of the importance of direct assessment of protein products comprising the basic functional units in cancer cells' biological processes. The announcement of the Cancer Proteome Atlas (\acrshort{tcpa}, http://tcpaportal.org) excited the cancer research community\cite{Li2013c}\cite{Li2017b}. By the utility of the \acrfull{rppa} or \acrfull{rpma}, a microarray of "Western blots" in the \acrshort{tcpa} could help to test our hypotheses.
However, there are only 237 antibodies available so far. We found that our candidates are not included in the \acrshort{tcpa} database (v3.0\cite{Chen2019}) yet. %  USP-10 in HPA
%> semi_join(pvalueTex_bad, HPA_bad, by=c("V1"="Gene"))
%validation at protein level:
%At the same time, we found our candidates (DKK1, CAMK2N1, STC2, PGK1, SURF4, NDFIP1, STIP1, DKC1) are also on the list of unfavorable prognostic genes for HNSCC from \acrfull{hpa} (available at https://www.proteinatlas.org/humanproteomepathology/head+and+neck+cancer, accessed 28 September 2020). The ZNF557, ZNF266, and FCGBP are on the list of favorable prognostic genes as well.
% > semi_join(pvalueTex_good, HPA_good, by=c("V1"="Gene"))
%There are 628 genes that consensus of pvalue005KM (6429 genes) and HPA (bad+good in HNSCC)
%%%%%%%%%% modified discussion
%\begin{MyColorPar}{red}

\subsection{our lab}
It is encouraged for multidisciplinary studies that use complementary computational and experimental approaches to address challenging cancer research.
%下一篇文章的內容。目前在計畫中，但還沒有開始進行
These in vitro and in vivo validation experiments will be undertaken in our laboratory.
We plan to analyze mRNA (e.x. qRT-PCR) and protein (e.x. western blot) of HNSCC cell lysate to confirm the candidate genes' expression.
%mRNA expression level in HNSCC cell lines (其他文章已有發表 overexpression）
The effect of over-expression and knockdown of the gene by lentiviral clones should be observed on cell function assay (e.x. proliferation, migration, and invasion) and mouse xenograft model (e.x. tumor growth). %這些都需要一年以上的時間
%their rationale for looking into TMSB4X is not strongly supported by the available data.

We utilized other approaches to validate our candidates.

\subsection*{Analysis by web-tools}
After discovering using our workflow, we also used the web-tools to find the evidence convincing up to 12 of our suggested 20 biomarkers in TCGA and GSE2837 cohort.
%\subsubsection*{SurvExpress} % TCGA
% 2021/01/23 pdf x 12 from SurvExpress
SurvExpress\cite{Aguirre-Gamboa2013} is one of the web-based tools for Biomarker comparison and validation of survival gene expression data. Using their TCGA HNSCC cohort (June 2016, n=502), the survival significance (Kaplan Meier \textit{P} value) shows in 12 genes:\\ % SurvExpress 有類似的結果:
1) poor prognosis: DKK1 (0.00005), CAMK2N1 (0.000006), STC2 (0.001), PGK1 (0.01), SURF4 (0.002), USP10 (0.002), NDFIP1 (0.000008), STIP1 (0.001), DKC1 (0.01);
%CAMK2N1 (0.048214), PGK1 (0.009978), SURF4 (0.023127), USP10 (0.017768), NDFIP1 (0.022758), FOXA2 (0.001587); % FOXA2 (0.038125)
2) better prognosis: ZNF557 (0.0007), ZNF266 (0.00008), FCGBP (0.001).
%IL19 (0.049731), FCGBP (0.005658), KIAA1683 (IQCN, 0.005886), NPB (0.014177);
%17 out of 20 candidates
%?DKK1 (0.253635),
The 12 out of 20 positive and negative prognostic genes underlined are confirmed by comparable studies using SurvExpress with the TCGA cohort. Please see Supplementary Figure S1. %\ref{fig:fig_SurvExpress}.
% moved to supplementary file

@@

%\subsubsection{HPA} %still TCGA's RNA-Seq
The Human Protein Atlas project (HPA) Proteome analysis is based on26941 antibodies targeting 17165 unique proteins. The HPA's Pathology Atlas analyzes each protein in patients, using \acrshort{ihc} analysis based on tissue microarrays (TMAs) adopted from TCGA. Kaplan-Meier survival analyses are based on RNA-Seq expression levels of human genes in HNSCC tissue and the clinical outcome.
All transcriptomics data has been retrieved from the Cancer Genome Atlas, and all proteomics data has been generated in-house using the same antibodies.
Our candidates (DKK1, CAMK2N1, STC2, PGK1, SURF4, USP10, NDFIP1, STIP1, DKC1) are also on the list of unfavorable prognostic genes for HNSCC from \acrfull{hpa} (available at https://www.proteinatlas.org/humanproteome/pathology/head+and+neck+cancer, Version: 20.0 updated: 2020-11-19). The ZNF557, ZNF266, and FCGBP are on the list of favorable prognostic genes as well.
Moreover, overexpression of ubiquitin-specific peptidase 10 (USP10), one of the autophagy-related genes, has a poor prognostic impact on HNSCC, which has been proved at mRNA and protein level((HPA database, available at https://www.proteinatlas.org/ENSG00000103194-USP10/pathology/head+and+neck+\\cancer)\cite{Ren2020}.
% 13 ARGs (GABARAPL1, ITGA3, USP10, ST13, MAPK9, PRKN, FADD, IKBKB, ITPR1, TP73, MAP2K7, CDKN2A, and EEF2K) with prognostic value were identified in HNSCC patients
%found at Human Protein Atlas dataset 
(Please see Supplementary Figure S2.) %\ref{fig:fig_HPA_USP10})
% moved to Supplementary

@@

%\subsubsection*{GSE2837} % non-TCGA
% from Anwser2-2
%by GEO GSE2837 HNSCC dataset (PrognoScan)
We analysed GSE2837 dataset (NCBI GEO database\cite{Chung2006}, HNSCC cohort has 28 participants) from online tools: PrognoScan (availalbe at http://dna00.\\bio.kyutech.ac.jp/PrognoScan/)\cite{Mizuno2009a}.
%??MD Anderson (40 cases) SurvNet: https://bioinformatics.mdanderson.org/SurvNet/
%https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE2837 \cite{Chung2006}
The GSE2837 carries % VUMC, VAMC, UTMDACC (1992-2005)
relapse-free survival (RFS) data, instead of overall survival in TCGA dataset,
and microarray gene expression (by Affymetrix X3P chips U133\_X3P).
The survival significance (Kaplan Meier \textit{P} value) is in the following probes of 10 genes:\\
1) poor prognosis: CAMK2N1 (0.048214), PGK1 (0.009978), SURF4 (0.023127), USP10 (0.017768), NDFIP1 (0.022758), FOXA2 (0.001587);\\ % FOXA2 (0.038125)
2) better prognosis: IL19 (0.049731), FCGBP (0.005658), KIAA1683 (IQCN, 0.005886), NPB (0.014177);
%17 out of 20 candidates
%?DKK1 (0.253635), 
Please see Supplementary Figure S3.\\ %\ref{figure:fig_GSE2837}.
Nevertheless, PrognoScan has group separation cut by a skewed manner, and the GSE2837 has far fewer participants than the TCGA cohort.
%DKK1, CAMK2N1, STC2, PGK1, SURF4, USP10, NDFIP1, FOXA2, STIP1, DKC1;
%ZNF557, ZNF266, IL19, MYO1H, FCGBP, LOC148709, EVPLL, PNMA5, KIAA1683 (IQCN), NPB
%Those 11 genes achieve similar positive and negative prognostic effect comparable with our proposed candidate genes. 
The 10 out of 20 positive and negative prognostic genes underlined is confirmed by comparable studies using the GSE2837 dataset other than the TCGA cohort.

\subsection*{Review articles}
After discovering using our workflow, we also performed the literature review by Embase/Pubmed to find the evidence convincing to 7 of our suggested 20 biomarkers in cancer research.
%Embase searching; %https://www.ncbi.nlm.nih.gov/research/pubtator/?view=docsum&query=CAMK2N1%20head%20and%20neck%20cancer
%through the PubMed searching, the remark on table 1 and table 2 presents the cancer research articles related to our candidate genes.
%PubTator Central\cite{Wei2019} is a useful Pubmed text miner

%%%%%% bad guy
The Dickkopf1 (DKK1) gene encodes a protein mainly involved in Wnt and other signaling pathways.
Inhibition of DKK1 in Hep-2 cells reduces their proliferation, colony formation, cell migration, and invasion in vitro\cite{Shi2014}.
Pang et al.\cite{Pang2018} demonstrated that upregulation of DKK1 in SBC-3 cells (human small cell lung cancer) enhanced their proliferation, colony formation, cell migration, and invasion in vitro, as well as bone metastasis in vivo. 
Increased DKK1 levels in HNSCC tissues correlated with elevated VEGF-C and beta-catenin\cite{Shi2014}.
DKK1 expression was significantly associated with smoking, alcohol abuse, sex, human papillomavirus status\cite{Chakraborty2020}, tumor site, tumor invasion, and pathologic stage in HNSCC patients.\cite{Gao2018}.
The mRNA expression of DKK1 and DKK3 was elevated in human papillomavirus (HPV)-negative HNSCC\cite{Hu2020}. Overexpression of DKK1 indicates adverse OS in bladder urothelial carcinoma (BLCA)\cite{Wei2020}, HNSCC\cite{Chakraborty2020}\cite{Hu2020}\cite{Wei2020}, and pancreatic adenocarcinoma (PAAD)\cite{Wei2020}. %Moreover, DKK1 is increased in HPV\+ HNSCC, leading to the worst prognosis of the patients. 
%\cite{Shi2014}\cite{Gao2018}\cite{Chakraborty2020}(non-TCGA)\cite{Hu2020}\cite{Wei2020}
%Wei et al\cite{Wei2020} conducted the analysis of prognostic value of DKK1 expression in human cancers based on bioinformatics tools, including UALCAN, GEPIA2 (dataset from TCGA)\cite{Tang2019}, and DriverDBv3 databases.
%using http://gepia2.cancer-pku.cn/detail.php?gene=DKK1 => \cite{Tang2019}
%GEPIA is a commonly used interactive website that plots expression profiles of given genes (from TCGA database)

%STC2:
The miR-381 suppresses cell proliferation, migration, and invasion in the HNSCC SCC-4 cell line through targeting stanniocalcin 2 (STC2) and participates in HNSCC development probably via the FAK/PI3K/Akt/mTOR signaling pathway.\cite{Ma2020}

% PGK1
Phosphoglycerate kinase 1 (PGK1 or PGK-1), a glycolysis enzyme, is responsive in cisplatin-resistant HNSCC cell line (H-1R). The resistance is associated with up-regulated expression of ATP-binding cassette transporter genes (MDR1, MRP1, and MRP2), CD55, and PGK1 and down-regulated expression Caveolin 1\cite{Nakamura2005}.

>>>>

\newenvironment{MyColorPar}[1]{%
    \leavevmode\color{#1}\ignorespaces%
}{%
}%
%\usepackage[upplemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
%...
%\usepackage{graphicx}
%\DeclareGraphicsExtensions{.pdf,.png} % for high-resolution PDF image
%\usepackage{hyperref}
\usepackage{xr}
% for xr in your preamble
% https://www.overleaf.com/learn/how-to/Cross_referencing_with_the_xr_package_in_Overleaf
\makeatletter
\newcommand*{\addFileDependency}[1]{% argument=file name and extension
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother

\newcommand*{\myexternaldocument}[1]{%
    \externaldocument{#1}%
    \addFileDependency{#1.tex}%
    \addFileDependency{#1.aux}%
}
% my external document I would like to reference the labels of. 
\myexternaldocument{Supplementary_figures} %.tex .aux

%\usepackage{url}

@@

%latex https://www.latex-tutorial.com/symbols/greek-alphabet/
%\subsection*{Linear regression in biomarker discovery}
Linear regression is useful as a concept in biomarker discovery.
A linear regression equation that relates a Y-variable to n X-variables is written as\\[1cm]
%\par
%Highlights
% https://quicklatex.com
\begin{flushleft}
% equation
%\begin{math}

@@

%\subsection{A introduction of TCGA}
%\begin{MyColorPar}{red}
%%%%%% revision1 added [TCGA] [2021/02/02]
A large-scale cancer database, aggregating many $X_1 ... X_n$ features, is necessary to facilitate the biomarker discovery.
The Cancer Genome Atlas (TCGA) project\cite{Weinstein2013} has been developed since 2005 and supervised by the National Cancer Institute's (NCI) Center for Cancer Genomics and the National Human Genome Research Institute (NHGRI), funded by the US government.
TCGA represents comprehensive genomics and clinic data from 84,392 patients among 33 major cancer types (data release 27.0 - Oct 29, 2020, available at https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga/studied-cancers).
%The Cancer Genome Atlas (TCGA): The TCGA consortium, which is a National Institute of Health (NIH) initiative, makes publicly available molecular and clinical information for more than 33 types of human cancers including exome (variant analysis), single nucleotide polymorphism (SNP), DNA methylation, transcriptome (mRNA), microRNA (miRNA) and proteome.
TCGA and other collaborated \acrfull{gdac} generated and analyzed DNA (mutation, copy number variation, methylation, simple nucleotide polymorphism, SNP), RNA (microarray, RNA-Seq, microRNA), and protein (reverse protein phased array) data derived from biospecimen. Sample types available at TCGA are primary solid tumors, recurrent solid tumors, blood-derived normal and tumor, metastatic, and solid tissue normal.

@@

% to summarize the advantages of its use.
The highlights of the advantages of applying the TCGA data for cancer biomarker identification include:
\begin{outline}
\1  To the best of our knowledge, the TCGA database is the largest collection (in both cancer types and cohort size) of comprehensive genomics with survival data available in the field of cancer research so far.  
There are many physical and social features ($X_1 ... X_n$) of patients available for survival modeling.
The whole-genome sequencing data has been harmonized across all GDACs.  Many databases adopt the essential demographic data from TCGA since it has comprehensive clinical features, such as exposure to alcohol, asbestos, radioactive radon, tobacco smoking, or cigarettes.
\1  TCGA has a remarkable advantage for computational and life scientists who study cancer since useful web-based tools and API are ready to analyze and visualize TCGA data. It might be getting help quickly from the research community for trouble-shooting.
\1  Many achievements of diagnoses, treatments, and prevention with the TCGA data have already been published and keep growing\cite{Tomczak2015}.\\[0.5cm]
\end{outline}

In summary, identifying predictive biomarkers for selecting standard-of-care or advanced systemic therapy\cite{Cristina2019} in \acrshort{hnscc} is crucial. 
However, there are three challenges of biomarker discovery from survival analysis.
First, although \acrshort{tcga} genomics data was harmonized, there is unclean data, including null expressed genes, over 50\% of the cohort, should be manually investigated and cleaned.
Second, we need to find a way to determine candidates from the expression level of 20,500 human protein-coding genes\cite{Clamp2007}. Usually, the investigators should get the rationale or revelation of the genes of interest on a specific cancer type. They should upload those genes manually onto bioinformatics tools, such as SurvExpress (http://bioinformatica.mty.itesm.mx:8080/Biomatec/ \newline
SurvivaX.jsp\cite{Aguirre-Gamboa2013}, which has been lost since Oct/2019 and currently out of funds), and analyze the cohort (e.x. \acrshort{tcga}). After downloading the survival results, they could curate plots and tables carefully.
It is not possible to scan the whole human protein-coding genome in this way.
Third, we need to find an optimal cutpoint of that \acrshort{rna} expression data to maximize candidate mining coverage. The above mentioned web-based tools might set a cutpoint at the median, 1/4 quantile, or 3/4 quantile for subsequent analyses. There are several visualization software or R packages which deal with cutoff determination, such as Prognoscan\cite{Mizuno2009a}, Cutoff Finder\cite{Budczies2012}, Findcut\cite{Chang2017a}, Human protein atlas\cite{Uhlen2017}, OptimalCutpoints\cite{Cristina2019},  cutpointr (available at https://github.com/thie1e/ \newline
cutpointr), and cutoffR (available at https://cran.r-project.org/web/
packages/cutoffR). However, non of them could combine the scanning of the protein-coding genes and cutoff optimization programmatically.

@@

%%%%
\subsection{Feature selection for survival modeling} % Even there are many $X_1 ... X_n$ physical and social features of patients available for survival modelling in the TCGA.
% TCGAbiolinks
%alcohol consumption
Besides ethnicity, age, gender, \acrshort{tnm} stage, radiation therapy, chemotherapy, and targeted therapy, the comprehensive adversely prognostic features in \acrshort{hnscc} should also include tobacco exposure, \acrshort{egfr} amplification, \acrfull{hpv} status, positive/close surgical margin ($<5 mm$), \acrfull{ene}, \acrfull{lvsi}, \acrfull{pni}, \acrfull{doi} ($>5 mm$), as well as metastatic \acrfull{lnd}\cite{Cheraghlou2018}, and \acrfull{wpoi5}, which is defined as tumor dispersion (1 mm apart between tumor satellites) or positive \acrshort{pni}/\acrshort{lvsi}\cite{Amin2017}.
The features of \acrshort{doi}, \acrshort{lnd}, and tumor dispersion are not available on the \acrshort{tcga} dataset. The Brandwein-Gensler's risk model (lymphocytic host response, \acrshort{wpoi5}, and \acrshort{pni})\cite{Brandwein-Gensler2010}\cite{Sinha2018} has been suggested to be routinely performed on pathological examination. In previous reports of \acrshort{hnscc}, the loco-regional failure will be high when the initial frozen section has a positive/close surgical margin, and even the final margin revision revealed negative\cite{Bulbul2019b}.
According to Table \ref{table:table2} and Table \ref{table:table4} in our study, the positive surgical margin yields a hazard ratio greater than 1.6 to influence on patient's \acrshort{os}.
It is suggested by authors \cite{Scholl1986}\cite{Sutton2003}\cite{Shaw2004}\cite{Guillemaud2010a}\cite{Patel2010}\cite{Kuriakose2017}\cite{Shapiro2017}\cite{Saidak2018}\cite{Miguelanez-Medran2019}\cite{Saidak2019} that the reason of positive/close surgical margin is possibly due to tumor aggressiveness or dispersion (WPOI-5) instead of iatrogenic reason of surgery. The surgical margin status has also been suggested as an independent surrogate for tumor dispersion in the \acrshort{hnscc} study. Thus, we selected common clinicopathological features in the current biomarker discovery, including gender, age, clinical T, clinical N, clinical M, surgical margin status, and tobacco exposure, to adjust confounders (details description at Materials and Methods section).

%%%

\subsection{Validation by web-based tools}
We utilized another approach of web-based tools (TCGA and GSE2837 cohort) to validate our candidates. %Since the discovery of workflow, we also used the web-based tools to find the evidence convincing up to 12 of our suggested 20 biomarkers in TCGA and GSE2837 cohort.

%\subsubsection*{SurvExpress} % TCGA
% 2021/01/23 pdf x 12 from SurvExpress
SurvExpress\cite{Aguirre-Gamboa2013} is one of the web-based tools for Biomarker comparison and validation of survival gene expression data. Using their TCGA HNSCC cohort (June 2016, n=502), the survival significance (Kaplan Meier \textit{P} value) shows in 12 genes:\\ % SurvExpress 有類似的結果:
1) poor prognosis: DKK1 (0.00005), CAMK2N1 (0.000006), STC2 (0.001), PGK1 (0.01), SURF4 (0.002), USP10 (0.002), NDFIP1 (0.000008), STIP1 (0.001), DKC1 (0.01);\\
%CAMK2N1 (0.048214), PGK1 (0.009978), SURF4 (0.023127), USP10 (0.017768), NDFIP1 (0.022758), FOXA2 (0.001587); % FOXA2 (0.038125)
2) better prognosis: ZNF557 (0.0007), ZNF266 (0.00008), FCGBP (0.001).\\
%IL19 (0.049731), FCGBP (0.005658), KIAA1683 (IQCN, 0.005886), NPB (0.014177);
%17 out of 20 candidates
%?DKK1 (0.253635),
The 12 out of 20 candidate genes are confirmed by a comparative study using SurvExpress with the TCGA cohort. Please see Supplementary Figure S1.%\ref{fig:fig_SurvExpress}.
% moved to supplementary file: Supplementary_figures.tex

@@

%\subsubsection{HPA} %still TCGA's RNA-Seq
The Human Protein Atlas project (HPA) has proteomics analysis based on 26,941 antibodies targeting 17,165 unique proteins. The HPA's Pathology Atlas analyzes each protein in patients, using \acrshort{ihc} analysis based on tissue microarrays (TMAs) adopted from TCGA. Kaplan-Meier survival analyses are based on RNA-Seq expression levels of human genes in HNSCC tissue and the clinical outcome.
All transcriptomics data has been retrieved from the TCGA, and all proteomics data has been generated in-house using the same antibodies.
Our 9 candidates (e.g., DKK1, CAMK2N1, STC2, PGK1, SURF4, USP10, NDFIP1, STIP1, and DKC1) are also on the list of unfavorable prognostic genes for HNSCC from \acrfull{hpa} (available at https://www.proteinatlas.org/humanproteome/pathology/head+and+neck+cancer, Version: 20.0 updated: 2020-11-19). The ZNF557, ZNF266, and FCGBP are on the list of favorable prognostic genes as well.
Moreover, overexpression of ubiquitin-specific peptidase 10 (USP10), one of the autophagy-related genes, has a poor prognostic impact on HNSCC, which has been proved at mRNA and protein level (HPA database, available at https://www.proteinatlas.org/ENSG00000103194-USP10/pathology/head+and+neck+\\cancer)\cite{Ren2020}.
% 13 ARGs (GABARAPL1, ITGA3, USP10, ST13, MAPK9, PRKN, FADD, IKBKB, ITPR1, TP73, MAP2K7, CDKN2A, and EEF2K) with prognostic value were identified in HNSCC patients
%found at Human Protein Atlas dataset 
(Please see Supplementary Figure S2)%\ref{fig:fig_HPA_USP10})
% moved to Supplementary

@@

%\subsubsection*{GSE2837} % non-TCGA
% from Anwser2-2
%by GEO GSE2837 HNSCC dataset (PrognoScan)
We analysed GSE2837 dataset (NCBI GEO database\cite{Chung2006}, HNSCC cohort has 28 participants) from online tools: PrognoScan (availalbe at http://dna00.bio.kyutech.ac.jp/\\PrognoScan/)\cite{Mizuno2009a}.
%??MD Anderson (40 cases) SurvNet: https://bioinformatics.mdanderson.org/SurvNet/
%https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE2837 \cite{Chung2006}
The GSE2837 carries % VUMC, VAMC, UTMDACC (1992-2005)
relapse-free survival (RFS) data, instead of overall survival in TCGA dataset,
and microarray gene expression (by Affymetrix X3P chips U133\_X3P).
The survival significance (Kaplan Meier \textit{P} value) is in the following probes of 10 genes:\\
1) poor prognosis: CAMK2N1 (0.048214), PGK1 (0.009978), SURF4 (0.023127), USP10 (0.017768), NDFIP1 (0.022758), FOXA2 (0.001587);\\ % FOXA2 (0.038125)
2) better prognosis: IL19 (0.049731), FCGBP (0.005658), KIAA1683 (IQCN, 0.005886), NPB (0.014177);\\
%17 out of 20 candidates
%?DKK1 (0.253635), 
Please see Supplementary Figure S3.%\ref{fig:fig_GSE2837}.

Nevertheless, PrognoScan has group separation cut by a skewed manner, and the GSE2837 has far fewer participants than the TCGA cohort.
%DKK1, CAMK2N1, STC2, PGK1, SURF4, USP10, NDFIP1, FOXA2, STIP1, DKC1;
%ZNF557, ZNF266, IL19, MYO1H, FCGBP, LOC148709, EVPLL, PNMA5, KIAA1683 (IQCN), NPB
%Those 11 genes achieve similar positive and negative prognostic effect comparable with our proposed candidate genes. 
The 10 out of 20 candidates are confirmed by a comparative study using the GSE2837 dataset other than the TCGA cohort.

%%%
\subsection{The candidates shown in literature review}
% *** why here?
In our previous work, altered glucose metabolism (e.k.a. the Warburg effect\cite{Warburg1956}) promotes the progression of \acrshort{hnscc}, which is partially attributed to the \acrfull{slc2a4} (or \acrlong{glut4}, \acrshort{glut4}) and \acrfull{trim24} pathway\cite{Chang2017b}.

%%
%\subsection*{Review articles}
Since the discovery of workflow, we performed the literature review by Embase/Pubmed.% to find the evidence convincing the suggested biomarkers in cancer research. 
The 10 out of 20 candidates has also been suggested by published studies using TCGA cohort, in-house cohort, or experiments in vitro and in vivo.

%Embase searching; %https://www.ncbi.nlm.nih.gov/research/pubtator/?view=docsum&query=CAMK2N1%20head%20and%20neck%20cancer
%through the PubMed searching, the remark on table 1 and table 2 presents the cancer research articles related to our candidate genes.
%PubTator Central\cite{Wei2019} is a useful Pubmed text miner

%%%%%% bad guy
The Dickkopf1 (DKK1) gene encodes a protein mainly involved in Wnt and other signaling pathways.
Inhibition of DKK1 in Hep-2 cells reduces their proliferation, colony formation, cell migration, and invasion in vitro\cite{Shi2014}.
Pang et al.\cite{Pang2018} has proved that upregulation of DKK1 in SBC-3 cells (human small cell lung cancer) enhances their proliferation, colony formation, cell migration, and invasion in vitro, as well as bone metastasis in vivo. 
Increased DKK1 levels in HNSCC tissues is correlated with elevated VEGF-C and beta-catenin\cite{Shi2014}.
DKK1 expression is significantly associated with smoking, alcohol abuse, human papillomavirus status\cite{Chakraborty2020}, tumor site, tumor invasion, and pathologic stage in HNSCC patients.\cite{Gao2018}.
The mRNA expression of DKK1 and DKK3 is elevated in human papillomavirus (HPV)-negative HNSCC\cite{Hu2020}. Overexpression of DKK1 indicates adverse OS in bladder urothelial carcinoma (BLCA)\cite{Wei2020}, HNSCC\cite{Chakraborty2020}\cite{Hu2020}\cite{Wei2020}, and pancreatic adenocarcinoma (PAAD)\cite{Wei2020}. %Moreover, DKK1 is increased in HPV\+ HNSCC, leading to the worst prognosis of the patients. 
%\cite{Shi2014}\cite{Gao2018}\cite{Chakraborty2020}(non-TCGA)\cite{Hu2020}\cite{Wei2020}
%Wei et al\cite{Wei2020} conducted the analysis of prognostic value of DKK1 expression in human cancers based on bioinformatics tools, including UALCAN, GEPIA2 (dataset from TCGA)\cite{Tang2019}, and DriverDBv3 databases.
%using http://gepia2.cancer-pku.cn/detail.php?gene=DKK1 => \cite{Tang2019}
%GEPIA is a commonly used interactive website that plots expression profiles of given genes (from TCGA database)

%STC2:
The miR-381 suppresses cell proliferation, migration, and invasion in the HNSCC SCC-4 cell line through targeting stanniocalcin 2 (STC2) and participates in HNSCC development probably via the FAK/PI3K/Akt/mTOR signaling pathway\cite{Ma2020}.

% PGK1
Phosphoglycerate kinase 1 (PGK1 or PGK-1), a glycolysis enzyme, is responsive in cisplatin-resistant HNSCC cell line (H-1R). 
The resistance is associated with the up-regulated expression of PGK1, CD55, ATP-binding cassette transporter genes (e.g., MDR1, MRP1, and MRP2), and down-regulated expression Caveolin 1\cite{Nakamura2005}.
